---
title: "2016 experiment"
author: "Alison Fowler"
date: "11/17/2020"
output: 
  html_document:
    toc: true
    toc_depth: 5
---

This R code is for analyzing the effect of pollen diet on Crithidia infection in wild Bombus griseocollis and Bombus bimaculatus workers for the manuscript titled: "Sunflower pollen reduces a gut pathogen in the model bee species, Bombus impatiens, but has weaker effects in three wild congeners"

Authors: Alison Fowler*, Jonathan Giacomini, Sara June Connon, Becky Irwin, and Lynn Adler 
Corresponding author*: aefowler@umass.edu 

## Libraries 

```{r, results = 'hide', message = FALSE, warning = FALSE}
library(lattice)
library(psych)
library(car)
library(AER)
library(MASS)
library(corrplot)
library(AICcmodavg)
library(plyr)
library(dplyr)
library(tidyr)
library(lme4)
library(mgcv)
library(scatterplot3d)
library(bbmle)
library(emmeans)
library(ggplot2)
library(tidyverse)
library(ggalt)
library(multcomp)
library(coxme) # for survival analyses with cox proportional hazards mixed model 
library(survival) # for survival analyses with survfit and Surv functions 
library(DHARMa)
library(aod) # wald test 
library(pscl) # interpreting binomial model output 
```

## Set up 

```{r, message = FALSE}
bees <- read_csv("CSV/2016_singles_all_spp.csv")
```

```{r}
# make treat a factor 
bees$Treatment<-as.factor(bees$Treatment)

# make spp a factor 
bees$species<-as.factor(bees$species)

# reorder species 
bees$species<- factor(bees$species, levels=c("Com imp", "Wild imp", "Wild gris"))

# rename levels of diet - just make new column actually 
bees <- bees %>% 
  mutate(Diet = grepl("S", bees$Treatment, fixed = T))
# right now, it says TRUE if it was S, and FALSE if it was B 
# now replace those words with our diets 
bees$Diet <- replace(bees$Diet, bees$Diet=="TRUE", "SF")
bees$Diet <- replace(bees$Diet, bees$Diet=="FALSE", "BW")

# reorder diets 
bees$Diet<- factor(bees$Diet, levels=c("SF", "BW"))

# rename species levels 
levels(bees$species)
bees$Species <- revalue(bees$species, c(
  "Com imp"="Commercial B. impatiens",
  "Wild imp"="Wild B. impatiens",
  "Wild gris"="Wild B. griseocollis"))
levels(bees$Species)
```

```{r}
# remove dead 
bees_alive <- bees %>%
  filter(!is.na(count))
```

## Count analysis, all three "species" together, zeros and non-zero counts all together 

### Data Exploration 

```{r}
summary(bees$Treatment)
summary(bees$species)

summary(bees_alive$Treatment)
summary(bees_alive$species)

boxplot(count ~ Treatment*species, data=bees_alive)

hist(bees_alive$count)

plot(count ~ wing, data=bees_alive)

boxplot(wing ~ Species, data = bees) # bee size correlated with species - griseocollis are bigger. 
plot(wing ~ count, data = bees)
```

### Model Selection 

```{r}
mod2016 <- glm.nb(count ~ Diet*Species + wing, data = bees)
mod2016_2 <- glm.nb(count ~ Diet*Species, data = bees)
mod2016_3 <- glm.nb(count ~ Diet + Species + wing, data = bees)
mod2016_4 <- glm.nb(count ~ Diet + Species, data = bees)

AICtab(mod2016, mod2016_2, mod2016_3, mod2016_4)
```

Model validation: 
```{r}
summary(mod2016)
par(mfrow=c(2,2))
plot(mod2016)
simresid <- simulateResiduals(mod2016, plot = T) # not great 
```

Lets look at model 2 
```{r}
simresid <- simulateResiduals(mod2016_2, plot = T) # better.... not perfect but better. 
# and excluding wing makes sense since it is confounded with species 
hist(mod2016_2$residuals)
```

```{r}
Anova(mod2016_2, type = "III")
```

## Logistic regression for probability of getting infected - all 3 bee groups together 

```{r}
# add a column for present/absent 

bees_alive$Presence <- NA

for(i in 1:length(bees_alive$count)){
  if(bees_alive$count[i] >=1) {
    bees_alive$Presence[i] <- 1
  } else {
    bees_alive$Presence[i] <- 0
  }
} 
```

Model 
```{r}
presence_mod <- glm(Presence ~ Diet * Species, family = binomial(link = "logit"), data = bees_alive)
```

```{r}
summary(presence_mod)
Anova(presence_mod)
```


## B. impatiens alone 

```{r}
imps16 <- bees_alive %>% 
  filter(species != "Wild gris")
```

### Model 

We're not going to include colony as a random effect because its confounded with 'species' - wild imps were all from the same colony. commercial imps came 3 colonies, which is less than 5 levels anyway so don't include. 

```{r}
hist(imps16$count)
imps.mod <- glm.nb(count ~ Diet*Species + wing, data = imps16)
imps.mod2 <- glm.nb(count ~ Diet + Species + wing, data = imps16)
imps.mod3 <- glm.nb(count ~ Diet * Species, data = imps16)
imps.mod4 <- glm.nb(count ~ Diet + Species, data = imps16)

AICtab(imps.mod, imps.mod2, imps.mod3, imps.mod4)
```

```{r}
simresid <- simulateResiduals(imps.mod3, plot = T)
```

```{r}
simresid <- simulateResiduals(imps.mod, plot = T)
```

```{r}
simresid <- simulateResiduals(imps.mod4, plot = T)
```

```{r}
simresid <- simulateResiduals(imps.mod2, plot = T)
```

```{r}
Anova(imps.mod3)
```

### Plot Impatiens 

```{r}
imp16_means<-emmeans(imps.mod3, ~Diet*Species, type = "response") 
imp.means.to.plot<-as.data.frame(summary(imp16_means))
imp.means.to.plot

imp.means.to.plot$mean<- imp.means.to.plot$response
imp.means.to.plot$upper<-imp.means.to.plot$mean + imp.means.to.plot$SE
imp.means.to.plot$lower<-imp.means.to.plot$mean - imp.means.to.plot$SE
imp.means.to.plot

imp16.lets<-cld(imp16_means, Letters = c('a','b','c', 'd'), sort = FALSE)
imp16.lets
imp16.letters<-dplyr::select(imp16.lets, Diet, .group)
imp16.letters

#mod2016.letters$letters <- factor(c(" a ", " b ", " b ", " c ", "bc", "bc")) # because the spacing is annoying 
#mod2016.letters
```

```{r}
imp_2016_plot<-
  ggplot(imp.means.to.plot, aes(x=Diet, group = Diet,
                          y=mean, 
                          ymin=lower,
                          ymax=upper)) + 
  theme_classic()+
  geom_point(stat="identity",
             aes(color = Diet,
                 shape = Species),
             size = 5) +
  geom_errorbar(width=0.05, aes(color = Diet))+
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"))+
  theme(text = element_text(size=15))+
  geom_text(aes (y = imp.means.to.plot$mean, 
                 label = imp16.letters$.group), 
            size = 5, hjust = 1.5) +
  scale_shape_manual("Origin",
                     labels = c("Commercial",
                                "Wild"),
                     values = c(15,17),
                     guide = "legend") + 
  scale_color_manual(values = c("chocolate2", "royalblue3")) + 
  guides(color = FALSE) +
  theme(legend.position = c(0.3, 0.75)) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.title.y = element_blank()) + 
  ggtitle(expression(paste(italic("B. impatiens")))) +
  geom_text(x = 2.35, y = 403, label = "2016", size = 4)

```

### Presence/absence 

Wild imps didn't have any bees with counts of 0. lets just look at commercial impatiens.

#### Look at just commercial bees 
```{r}
comm_imps <- imps16 %>% 
  filter(Species == "Commercial B. impatiens")

comm_imps_presence_mod <- glm(Presence ~ Diet, family = binomial(link = "logit"), data = comm_imps)

comm_imps_presence_mod2 <- glm(Presence ~ Diet + wing, family = binomial(link = "logit"), data = comm_imps)

AICtab(comm_imps_presence_mod, comm_imps_presence_mod2)

summary(comm_imps_presence_mod)
Anova(comm_imps_presence_mod)
```

```{r}
Diet <- c("BW", "SF")
Species <- c("Commercial B. impatiens","Commercial B. impatiens")
newdata_comm <- data.frame(Species, Diet)
newdata_comm$dietP <- predict(comm_imps_presence_mod, newdata = newdata_comm, type = "response")
newdata_comm
```

```{r}
newdata_comm_2 <- cbind(newdata_comm, predict(comm_imps_presence_mod, newdata = newdata_comm, type = "link", se = T))
newdata_comm_2
```

```{r}
newdata_comm_3 <- within(newdata_comm_2, {
    PredictedProb <- plogis(fit)
    LL <- plogis(fit - (1.96 * se.fit)) # this translates to a 95% CI  
    UL <- plogis(fit + (1.96 * se.fit))
})

newdata_comm_3
```

## B. griseocollis alone 

```{r}
gris16 <- bees_alive %>% 
  filter(species == "Wild gris")
```

### Model 

```{r}
hist(gris16$count)
gris.mod <- glm.nb(count ~ Diet + wing, data = gris16)
gris.mod2 <- glm.nb(count ~ Diet, data = gris16)

AICtab(gris.mod, gris.mod2)
```

```{r}
plot(gris.mod)
simresid <- simulateResiduals(gris.mod, plot = T)
```

```{r}
Anova(gris.mod)
```

### Plot Griseocollis 

```{r}
gris16_means<-emmeans(gris.mod, ~Diet, type = "response") 
gris.means.to.plot<-as.data.frame(summary(gris16_means))
gris.means.to.plot

gris.means.to.plot$mean<- gris.means.to.plot$response
gris.means.to.plot$upper<-gris.means.to.plot$mean + gris.means.to.plot$SE
gris.means.to.plot$lower<-gris.means.to.plot$mean - gris.means.to.plot$SE
gris.means.to.plot

gris16.lets<-cld(gris16_means, Letters = c('a','b','c', 'd'), sort = FALSE)
gris16.lets
gris16.letters<-dplyr::select(gris16.lets, Diet, .group)
gris16.letters

#mod2016.letters$letters <- factor(c(" a ", " b ", " b ", " c ", "bc", "bc")) # because the spacing is annoying 
#mod2016.letters
```

PLOT FOR GRIS: 

```{r}
gris_2016_plot<-
  ggplot(gris.means.to.plot, aes(x=Diet, group = Diet,
                          y=mean, 
                          ymin=lower,
                          ymax=upper)) + 
  theme_classic()+
  geom_point(stat="identity",
             aes(color = Diet),
             size = 5) +
  geom_errorbar(width=0.05, aes(color = Diet))+
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"))+
  theme(text = element_text(size=15))+
  scale_color_manual(values = c("chocolate2", "royalblue3")) + 
  guides(color = FALSE) +
  theme(legend.position = "right") + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.title.y = element_blank()) + 
  ggtitle(expression(paste(italic("B. griseocollis ")))) + 
  geom_text(x = 2.35, y = 357.5, label = "2016", size = 4) 
#  theme(plot.title = element_text(hjust = 0.5))
```

### Presence/Absence 

```{r}
gris_presence_mod <- glm(Presence ~ Diet, family = binomial(link = "logit"), data = gris16)
gris_presence_mod2 <- glm(Presence ~ Diet + wing, family = binomial(link = "logit"), data = gris16)
AICtab(gris_presence_mod, gris_presence_mod2)
```

```{r}
summary(gris_presence_mod)
Anova(gris_presence_mod)
```

## Mortality 

```{r}
# make a survival object
event <- bees$Dead
time<-bees$Days_till_dead
survival<-Surv(time, event)
```

```{r}
# model survival as the response 
# can't include wing because dead bees don't have wing measurements 

mort1<- coxme(survival ~ species*Treatment + (1|colony), data=bees) 
mort2<- coxme(survival ~ species + Treatment + (1|colony), data=bees) 
mort3<- coxph(survival ~ species + Treatment, data=bees) 

AICtab(mort1, mort2, mort3)
```

```{r, eval= FALSE}
Anova(mort3)
```

