---
title: "2017 Experiment"
output: 
  html_document:
    toc: true
    toc_depth: 6
    highlight: pygments
---
This R code is for analyzing the effect of pollen diet on Crithidia infection in wild Bombus griseocollis and Bombus bimaculatus workers for the manuscript titled: "Sunflower pollen reduces a gut pathogen in the model bee species, Bombus impatiens, but has weaker effects in three wild congeners"

Authors: Alison Fowler*, Jonathan Giacomini, Sara June Connon, Becky Irwin, and Lynn Adler 
Corresponding author*: aefowler@umass.edu 

# Notes
Crithidia strain names: 
IMP-MA-SS is IMP-2
MOA is IMP-3 
GRIS-NC-JCRA is GRIS

# Libraries 
```{r, results = 'hide', message = FALSE, warning = FALSE}
library(lattice)
library(psych)
library(car)
library(AER)
library(MASS)
library(corrplot)
library(AICcmodavg)
library(dplyr)
library(tidyr)
library(lme4)
library(mgcv)
library(scatterplot3d)
library(bbmle)
library(emmeans)
library(ggplot2)
library(tidyverse)
library(ggalt)
library(multcomp)
library(coxme)
library(survminer)
library(DHARMa)
library(glmmTMB)
library(plyr)
```



# Look at both species together: 

## Get set up 

```{r, results = 'hide', message = FALSE}
bees2017<-read_csv("CSV/2017_gris_bimacs_final.csv")
```

```{r}
# rename columns
bees2017 <- bees2017 %>% 
  dplyr::rename(count = "Crithidia Count") %>% 
  dplyr::rename(ID = "Bee ID") %>% 
  dplyr::rename(inoc_date = "Inoc. Date") %>% 
  dplyr::rename(crithidia_strain = "Crithidia Strain") %>% 
  dplyr::rename(colony = "Colony ID")

# make species a factor
bees2017$Species <- as.factor(bees2017$Species)

# make treatment a factor
bees2017$Treatment <- as.factor(bees2017$Treatment)

# make colony ID a factor 
bees2017$colony<-as.factor(bees2017$colony)

# make date a factor 
bees2017$inoc_date<-as.factor(bees2017$inoc_date)

# make strain a factor
bees2017$crithidia_strain <- as.factor(bees2017$crithidia_strain)

# rename levels
levels(bees2017$Treatment)

bees2017$Treatment <- revalue(bees2017$Treatment, c(
  "50/50"="50-50",
  "CS100"="SF",
  "KW100"="WF"))

# reorder treatments 
bees2017$Treatment <- factor(bees2017$Treatment, levels=c("SF", "50-50", "WF"))
levels(bees2017$Treatment)

# make count and wing numeric 
bees2017$count <- as.numeric(paste(bees2017$count))
bees2017$MCL <- as.numeric(paste(bees2017$MCL))
```

### Clean up the data: 

```{r}
# select just workers
bees2017<- bees2017 %>% 
  filter(Caste == "worker")  

# remove unusable data points
bees2017 <- bees2017 %>% 
  filter(ID != "268") %>% # remove bees that died before inoculated
  filter(ID != "277") %>% 
  filter(ID != "211") %>%  # missing bee 
  filter(ID != "47") %>% # remove bees that didn't finish their inoculum
  filter(ID != "263") %>% 
  filter(ID != "282") %>% 
  filter(ID != "286") %>% 
  filter(ID != "287") %>% 
  filter(ID != "214")

# remove dead 
bees2017_alive <- bees2017 %>% 
  filter(count != ".")

# add a column for present/absent 
bees2017_alive$Presence <- NA

for(i in 1:length(bees2017_alive$count)){
  if(bees2017_alive$count[i] >=1) {
    bees2017_alive$Presence[i] <- 1
  } else {
    bees2017_alive$Presence[i] <- 0
  }
} 
```

Sample size including dead bees: 190 (69 bimacs & 121 gris)
69 out of 190 bees died (17 bimacs & 52 gris)
Sample size excluding dead bees: 121 (52 bimacs & 69 gris)

## Explore full data set 
```{r}
summary(bees2017$Treatment)
summary(bees2017$Species)
summary(bees2017$count)
```

```{r}
summary(bees2017_alive$Treatment)
summary(bees2017_alive$Species)
summary(bees2017_alive$count)
```
alive/total
Bimacs: 52/69
Gris: 69/121

## Data exploration with SS Crithidia 

```{r}
# lets subset out bees infected with SS Crithidia 
bees2017_SS <- bees2017 %>% 
  filter(crithidia_strain == "IMP-MA-SS")
```

Sample size of bees fed SS Critihdia: 138

```{r}
summary(bees2017_SS$Species) #45 bimacs, 93 gris 
summary(bees2017_SS$inoc_date)
summary(bees2017_SS$colony)
summary(bees2017_SS$Treatment)
```

```{r}
# remove dead bees
bees2017_SS_alive <- bees2017_SS %>% 
  filter(count != ".")

summary(bees2017_SS_alive$Species)
summary(bees2017_SS_alive$inoc_date)
summary(bees2017_SS_alive$colony)
summary(bees2017_SS_alive$Treatment)
```

Number of bees (out of 138) fed SS that survived the trial: 91 (66%)
Bimacs: 32/45 (71%)
Gris: 59/93 (63%)

```{r}
bees2017_SS_alive %>% 
  ggplot(aes(x = Treatment, y = count)) + 
  geom_boxplot() +
  theme_bw() + 
  geom_point(aes(color = Species), alpha = 0.5) + 
  scale_color_manual(values=c("darkorange", "deepskyblue", "darkred")) + 
  geom_encircle(aes(color=Species, group=Species), size = 2, lty = 2)

hist(bees2017_SS_alive$count)

plot(count ~ MCL, data = bees2017_SS_alive)

# replace lost wing values (NAs) with mean wing 
bees2017_SS_alive<- bees2017_SS_alive %>% 
  mutate(wing = case_when(is.na(MCL) ~ mean(MCL, na.rm = TRUE), 
                          TRUE ~ as.numeric(MCL)
                          )
         )
```

## Model both species infected with SS

```{r}
SSmod <- glmer.nb(count ~ Treatment*Species + wing + (1|inoc_date), data = bees2017_SS_alive)
isSingular(SSmod)
plot(SSmod)
Anova(SSmod)

SSmod_no_inoc<- glm.nb(count ~ Treatment*Species + wing, data = bees2017_SS_alive)
par(mfrow=c(2,2))
plot(SSmod_no_inoc)
Anova(SSmod_no_inoc)

SSmod_3<- glm.nb(count ~ Treatment*Species, data = bees2017_SS_alive)

AICtab(SSmod, SSmod_no_inoc, SSmod_3)

Anova(SSmod_3)
```

```{r}
SS_b_g_mod1 <- glmer.nb(count ~ Treatment*Species + wing + (1|inoc_date), data = SS_b_g)

SS_b_g_mod2 <- glm.nb(count ~ Treatment*Species + wing, data = SS_b_g)

SS_b_g_mod3 <- glm.nb(count ~ Treatment*Species, data = SS_b_g)

SS_b_g_mod4 <- glm.nb(count ~ Treatment + Species, data = SS_b_g)

AICtab(SS_b_g_mod1, SS_b_g_mod2, SS_b_g_mod3, SS_b_g_mod4)
```

```{r}
plot(SS_b_g_mod3)
summary(SS_b_g_mod3)
Anova(SS_b_g_mod3)
```

## Data exploration with MOA Crithidia 

```{r}
# lets subset out bees infected with MOA Crithidia 
bees2017_MOA <- bees2017 %>% 
  filter(crithidia_strain == "MOA")
```

Wow, a LOT of these bees either died or didnt have a count when dissected.

2 died before inoculation (268, 277), 14 died during the trial, 3 didn't drink their inoculum, and then 19 didn't have a count but didn't have a note that they died. 

So that leaves us with 23 bees, 20 of which are bimaculatus. 

```{r}
summary(bees2017_MOA$Species)
summary(bees2017_MOA$inoc_date)
summary(bees2017_MOA$colony)
summary(bees2017_MOA$Treatment)

bees2017_MOA %>% 
  ggplot(aes(x = Treatment, y = count)) + 
  geom_boxplot() +
  theme_bw() + 
  geom_point(aes(color = Species), alpha = 0.5) + 
  scale_color_manual(values=c("darkorange", "deepskyblue")) + 
  geom_encircle(aes(color=Species, group=Species), size = 2, lty = 2)

hist(bees2017_MOA$count)

plot(count ~ MCL, data = bees2017_MOA)
```

## Model Crithidia counts with both species and all strains 

```{r}
# treatment, species, strain
# interaction
b_g_mod <- glmer.nb(count ~ Treatment*Species*crithidia_strain + wing + (1|inoc_date), data = b_g_alive)
b_g_mod2 <- glm.nb(count ~ Treatment*Species*crithidia_strain + wing, data = b_g_alive)
b_g_mod3 <- glm.nb(count ~ Treatment*Species*crithidia_strain, data = b_g_alive)
#additive
b_g_mod4 <- glmer.nb(count ~ Treatment + Species + crithidia_strain + wing + (1|inoc_date), data = b_g_alive)
b_g_mod5 <- glm.nb(count ~ Treatment + Species + crithidia_strain + wing, data = b_g_alive)
b_g_mod6 <- glm.nb(count ~ Treatment + Species + crithidia_strain, data = b_g_alive)

# no strain 
#interaction
b_g_mod7 <- glmer.nb(count ~ Treatment*Species + wing + (1|inoc_date), data = b_g_alive)
b_g_mod8 <- glm.nb(count ~ Treatment*Species + wing, data = b_g_alive)
b_g_mod9 <- glm.nb(count ~ Treatment*Species, data = b_g_alive)
#additive
b_g_mod10 <- glmer.nb(count ~ Treatment + Species + wing + (1|inoc_date), data = b_g_alive)
b_g_mod11 <- glm.nb(count ~ Treatment + Species + wing, data = b_g_alive)
b_g_mod12 <- glm.nb(count ~ Treatment + Species, data = b_g_alive)

AICtab(b_g_mod, b_g_mod2, b_g_mod3, b_g_mod4, b_g_mod5, b_g_mod6, b_g_mod7, b_g_mod8, b_g_mod9, b_g_mod10, b_g_mod11, b_g_mod12)
```

```{r}
plot(b_g_mod11)
summary(b_g_mod11)
Anova(b_g_mod11)
```


# Analyze each species separately 

## B. bimaculatus 

### Get set up 

```{r, results = 'hide', message = FALSE}
bimacs<- bees2017 %>% 
  filter(Species == "B. bimaculatus")

summary(bimacs$crithidia_strain)
```

```{r}
bimacs_alive<- bees2017_alive %>% 
  filter(Species == "B. bimaculatus")
```

### Data exploration

```{r}
summary(bimacs$crithidia_strain) #total: 45 for SS and 34 for MOA
summary(bimacs_alive$crithidia_strain) #survived: 32/45 for SS and 20/34 for MOA 
summary(bimacs$inoc_date)
summary(bimacs$colony)
summary(bimacs$Treatment)

# colored by crithidia
bimacs %>% 
  ggplot(aes(x = Treatment, y = count)) + 
  geom_boxplot() +
  theme_bw() + 
  geom_point(aes(color = crithidia_strain), alpha = 0.5) + 
  scale_color_manual(values=c("deepskyblue", "darkred")) + 
  geom_encircle(aes(color=crithidia_strain, group=crithidia_strain), size = 2, lty = 2)

plot(count ~ MCL, data = bimacs)
hist(bimacs$count)
```

### Model crithidia counts 

```{r}
# all FEs and both REs
bimac_global <- glmer.nb(count ~ Treatment + crithidia_strain + MCL + (1|inoc_date) + (1|colony), data=bimacs)
bimac_global_int <- glmer.nb(count ~ Treatment * crithidia_strain + MCL + (1|inoc_date) + (1|colony), data=bimacs)

# all FEs and one RE
bimac_1 <- glmer.nb(count ~ Treatment + crithidia_strain + MCL + (1|colony), data=bimacs)
bimac_1_int <- glmer.nb(count ~ Treatment * crithidia_strain + MCL + (1|colony), data=bimacs)
bimac_2 <- glmer.nb(count ~ Treatment + crithidia_strain + MCL + (1|inoc_date), data=bimacs)
bimac_2_int <- glmer.nb(count ~ Treatment * crithidia_strain + MCL + (1|inoc_date), data=bimacs)

# ^ All of the above models gave warning for singular fit 

# all FEs and no REs 
bimac_3 <- glm.nb(count ~ Treatment + crithidia_strain + MCL, data=bimacs)
bimac_3_int <- glm.nb(count ~ Treatment * crithidia_strain + MCL, data=bimacs)

AICtab(bimac_global, bimac_global_int, bimac_1, bimac_1_int, bimac_2, bimac_2_int, bimac_3, bimac_3_int)
```

Model 3 with interaction and no REs does best 

```{r}
# now start removing FEs from bimac_3_int model 

# no wing 
bimac_4 <- glm.nb(count ~ Treatment * crithidia_strain, data = bimacs)

# no wing or strain 
bimac_5 <- glm.nb(count ~ Treatment, data = bimacs)

# no strain 
bimac_6 <- glm.nb(count ~ Treatment + MCL, data = bimacs)

AICtab(bimac_3_int, bimac_4, bimac_5, bimac_6)
```

MODEL TO USE for 2017 BIMACS: 
Model 4: count ~ Treatment * crithidia_strain

```{r}
par(mfrow=c(2,2))
plot(bimac_4)
simresid <- simulateResiduals(bimac_4, plot = T)
Anova(bimac_4)
summary(bimac_4)
```

### Plot

```{r}
bimac_means<-emmeans(bimac_4, ~Treatment*crithidia_strain) 
means.to.plot<-as.data.frame(summary(bimac_means))
means.to.plot$expmean<- exp(means.to.plot$emmean)
means.to.plot$tfupper<-exp(means.to.plot$emmean + means.to.plot$SE)
means.to.plot$tlower<-exp(means.to.plot$emmean - means.to.plot$SE)
exp.mean<-means.to.plot$expmean
means.to.plot

bimac.lets<-cld(bimac_means, Letters = c('a','b','c', 'd'), sort = FALSE)
bimac.letters<-dplyr::select(bimac.lets, Treatment, crithidia_strain, .group)
```

```{r}
bim_2017_plot<-
  ggplot(means.to.plot, aes(x=Treatment, group = Treatment,
                          y=means.to.plot$expmean, 
                          ymin=means.to.plot$tlower,
                          ymax=means.to.plot$tfupper)) + 
  theme_classic()+
  geom_point(stat="identity",
             aes(shape = crithidia_strain,
                 color = Treatment),
             size = 5) +
  geom_errorbar(width=0.05, aes(color = Treatment))+
  theme(axis.text.y = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"),
        axis.text.x = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"),
        text = element_text(size=18))+
  geom_text(aes (y = means.to.plot$expmean, 
                 label = bimac.letters$.group), 
            size = 5, hjust = 1.5) +
  scale_shape_manual(expression(paste(italic("Crithidia"))),
                     labels = c("IMP-2",
                                "IMP-3"),
                     values = c(15, 17),
                     guide = "legend") + 
  scale_x_discrete(labels = c("100% SF", "50% SF", "Control")) +
  scale_color_manual(name = "Pollen Diet",
                       labels = c("100% Sunflower",
                                  "50% Sunflower",
                                  "Control"),
                       values = c("chocolate2", "goldenrod1", "royalblue3")) +
 # guides(color = FALSE) +
  theme(legend.position=c(.2,.75)) + 
  theme(axis.title.x = element_blank()) +  
  theme(axis.title.y = element_blank()) + 
  geom_text(x = 3.35, y = 90, label = "2017", size = 4) + 
  ggtitle("")
  
```

### Crithidia presence absence for bimaculatus - only 4 out of 52 bimacs had counts of 0 

```{r}
# both RE 
bimac_presence1 <- glmer(Presence ~ Treatment + crithidia_strain + MCL + (1|colony) + (1|inoc_date), family = binomial(link = "logit"), data = bimacs_alive)

bimac_presence2 <- glmer(Presence ~ Treatment * crithidia_strain + MCL + (1|colony) + (1|inoc_date), family = binomial(link = "logit"), data = bimacs_alive)

# just colony RE
bimac_presence3 <- glmer(Presence ~ Treatment + crithidia_strain + MCL + (1|colony), family = binomial(link = "logit"), data = bimacs_alive)

bimac_presence4 <- glmer(Presence ~ Treatment * crithidia_strain + MCL + (1|colony), family = binomial(link = "logit"), data = bimacs_alive)

# just inoc date RE
bimac_presence5 <- glmer(Presence ~ Treatment + crithidia_strain + MCL + (1|inoc_date), family = binomial(link = "logit"), data = bimacs_alive)

bimac_presence6 <- glmer(Presence ~ Treatment * crithidia_strain + MCL + (1|inoc_date), family = binomial(link = "logit"), data = bimacs_alive)

#neither RE
bimac_presence7 <- glm(Presence ~ Treatment + crithidia_strain + MCL, family = binomial(link = "logit"), data = bimacs_alive)

bimac_presence8 <- glm(Presence ~ Treatment * crithidia_strain + MCL, family = binomial(link = "logit"), data = bimacs_alive)

AICtab(bimac_presence1, bimac_presence2, bimac_presence3, bimac_presence4, bimac_presence5, bimac_presence6, bimac_presence7, bimac_presence8)
```

7 is best out of these options 

```{r}
bimac_presence9 <- glm(Presence ~ Treatment + crithidia_strain, family= binomial(link = "logit"), data = bimacs_alive)

bimac_presence10 <- glm(Presence ~ Treatment, family= binomial(link = "logit"), data = bimacs_alive)

AICtab(bimac_presence7, bimac_presence9, bimac_presence10)
```

9 is best: presence ~ Treatment + strain 

``` {r}
summary(bimac_presence9)
Anova(bimac_presence9)
anova(bimac_presence9, test = "Chisq")
wald.test(b = coef(bimac_presence9), Sigma = vcov(bimac_presence9), Terms = 2) # term 2 = Diet 
```

```{r}
Treatment <- c("SF", "50-50", "WF")
crithidia_strain <- c("IMP-MA-SS","IMP-MA-SS", "IMP-MA-SS",
            "MOA", "MOA", "MOA")
newdata <- data.frame(crithidia_strain, Treatment)
newdata$dietP <- predict(bimac_presence9, newdata = newdata, type = "response")
newdata
```

```{r}
newdata2 <- cbind(newdata, predict(bimac_presence9, newdata = newdata, type = "link", se = T))
newdata2
```

```{r}
newdata3 <- within(newdata2, {
    PredictedProb <- plogis(fit)
    LL <- plogis(fit - (1.96 * se.fit)) # this translates to a 95% CI  
    UL <- plogis(fit + (1.96 * se.fit))
})

newdata3
```

Actually, we're only going to use IMP-SS subset to estimate presence since all MOA bees were infected 

```{r}
bimacs_17_SS <- bimacs_alive %>% 
  filter(crithidia_strain == "IMP-MA-SS")

bimac_presence11 <- glm(Presence ~ Treatment, family = binomial(link="logit"), data = bimacs_17_SS)
Anova(bimac_presence11)
simresid <- simulateResiduals(bimac_presence11, plot = T)
```

### Mortality 

```{r}
# make a survival object
event <- bimacs$dead
time <- bimacs$days_til_dead
survival <- Surv(time, event)
```

```{r}
# model survival as the response with the cox proportional hazards model 
mortb1<- coxph(survival ~ Treatment * crithidia_strain + MCL, data=bimacs) 
#Loglik converged before variable  7 ; coefficient may be infinite. 
mortb2<- coxph(survival ~ Treatment + crithidia_strain + MCL, data=bimacs) 

# both 3 and 4 received the following error: 
#Error in coxme.fit(X, Y, strats, offset, init, control, weights = weights, : No starting estimate was successful
#mortb3<- coxme(survival ~ Treatment * crithidia_strain + MCL + (1|inoc_date), data=bimacs)
mortb4<- coxme(survival ~ Treatment + crithidia_strain + MCL + (1|inoc_date), data=bimacs)

mortb5<- coxph(survival ~ Treatment + MCL, data=bimacs)
mortb6<- coxph(survival ~ crithidia_strain + MCL, data=bimacs)

mortb7<- coxph(survival ~ Treatment, data=bimacs)
mortb8<- coxph(survival ~ crithidia_strain, data=bimacs)

mortb9<- coxme(survival ~ Treatment + MCL + (1|inoc_date), data=bimacs)
#mortb10<- coxme(survival ~ crithidia_strain + MCL + (1|inoc_date), data=bimacs)
# 10 also received the same error 3 and 4 did 

mortb11<-coxph(survival ~ Treatment + crithidia_strain + MCL + inoc_date, data = bimacs)
#Loglik converged before variable  7 ; coefficient may be infinite. 

AICtab(mortb1, mortb2, mortb5, mortb6, mortb7, mortb8, mortb9)
```

```{r}
summary(mortb9)
Anova(mortb9)
summary(glht(mortb9,linfct=mcp(Treatment="Tukey")))
```


## B. griseocollis 

### Get set up 

```{r, results = 'hide', message = FALSE}
gris <- bees2017 %>% 
  filter(Species == "B. griseocollis")

summary(gris$crithidia_strain)
```

```{r}
gris_alive <- bees2017_alive %>% 
  filter(Species == "B. griseocollis")
```

```{r}
# remove the two other crithidia sources 
gris2<- gris %>% 
  filter(crithidia_strain == "IMP-MA-SS")
```

### Data exploration 

```{r}
summary(gris_alive)
summary(gris_alive$inoc_date)
summary(gris_alive$colony)
summary(gris_alive$Treatment)

# colored by crithidia
gris_alive %>% 
  ggplot(aes(x = Treatment, y = count)) + 
  geom_boxplot() +
  theme_bw() + 
  geom_point(aes(color = crithidia_strain), alpha = 0.5) + 
  scale_color_manual(values=c("deepskyblue", "darkred","violet")) + 
  geom_encircle(aes(color=crithidia_strain, group=crithidia_strain), size = 2, lty = 2)

# since there ended up being so few survivors in the other two crithidia source groups, I think we should analyze gris with just the SS. 

plot(count ~ MCL, data = gris2) + abline(a = exp(0.3420),
                                              b = 25.9)
hist(gris_alive$count)
```

```{r}
gris2 %>% 
  ggplot(aes(x = Treatment, y = count)) + 
  geom_boxplot() +
  theme_bw() + 
  geom_point(aes(color = colony), alpha = 0.5) + 
  scale_color_manual(values=c("deepskyblue", "darkred","violet", "darkorange", "gold")) + 
  geom_encircle(aes(color=colony, group=colony), size = 2, lty = 2)
```

### Model crithidia counts for just SS infected griseocollis 

```{r}
gris_global <- glmer.nb(count ~ Treatment + MCL + (1|inoc_date) + (1|colony), data=gris2)
gris_1 <- glmer.nb(count ~ Treatment +  MCL + (1|colony), data=gris2)
gris_2 <- glmer.nb(count ~ Treatment + MCL + (1|inoc_date), data=gris2)
isSingular(gris_2)
gris_3 <- glm.nb(count ~ Treatment + MCL, data=gris2)
gris_4 <- glm.nb(count ~ Treatment, data=gris2)

AICtab(gris_global, gris_1, gris_2, gris_3, gris_4)
```

```{r}
plot(gris_1)
simresid <- simulateResiduals(gris_1, plot = T) 
# deviates significantly - need to use another model 

plot(gris_global)
simresid <- simulateResiduals(gris_global, plot = T)
# still pretty bad 

plot(gris_3)
simresid <- simulateResiduals(gris_3, plot = T)

# problem: gris_global and gris_1 have lower AIC but don't pass simulation test. 
# gris_3 has higher AIC but does pass simulation test. 
# should we just use gris_3 then? 
```

```{r}
summary(gris_1)
hist(resid(gris_1))
```

```{r}
summary(gris_3)
hist(resid(gris_3))
```

So... why do the residuals look much more normal in gris_1 than gris_3, yet the simulation test seems to not reflect that? 

Not sure which model to use here, but let's go with the one that passes the DHARMA sim test: gris 3. 

```{r}
Anova(gris_1)
Anova(gris_3)
```

### Plot

```{r}
gris_means<-emmeans(gris_3, ~Treatment) 
g.17.means.to.plot<-as.data.frame(summary(gris_means))
g.17.means.to.plot$expmean<- exp(g.17.means.to.plot$emmean)
g.17.means.to.plot$tfupper<-exp(g.17.means.to.plot$emmean + g.17.means.to.plot$SE)
g.17.means.to.plot$tlower<-exp(g.17.means.to.plot$emmean - g.17.means.to.plot$SE)
exp.mean<-g.17.means.to.plot$expmean
g.17.means.to.plot

gris.lets<-cld(gris_means, Letters = c('a','b','c', 'd'), sort = FALSE)
gris.letters<-dplyr::select(gris.lets, Treatment, .group)
```

```{r}
gris_2017_plot<-
  ggplot(g.17.means.to.plot, aes(x=Treatment, group = Treatment,
                          y=g.17.means.to.plot$expmean, 
                          ymin=g.17.means.to.plot$tlower,
                          ymax=g.17.means.to.plot$tfupper)) + 
  theme_classic()+
  geom_point(stat="identity",
             aes(color = Treatment),
             size = 5) +
  geom_errorbar(width=0.05, aes(color = Treatment))+
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"))+
  theme(text = element_text(size=15))+
  scale_shape_manual(expression(paste(italic("Crithidia"))),
                     labels = c("IMP-2"),
                     values = c(15),
                     guide = "legend") + 
  scale_color_manual(values = c("chocolate2", "goldenrod1", "royalblue3")) + 
  guides(color = FALSE) +
  theme(legend.position = "right") + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.title.y = element_blank()) + 
  geom_text(x = 3.35, y = 110, label = "2017", size = 4) + 
  ggtitle("")
```

### Crithidia presence absence for gris - 15 out of 69 had counts of zero

```{r}
gris_presence <- glm(Presence ~ Treatment + MCL, family = binomial(link = "logit"), data = gris_alive)
gris_presence2 <- glmer(Presence ~ Treatment + MCL + (1|colony), family = binomial(link = "logit"), data = gris_alive) 
gris_presence3 <- glmer(Presence ~ Treatment + MCL + (1|colony) + (1|inoc_date), family = binomial(link = "logit"), data = gris_alive) 
gris_presence4 <- glmer(Presence ~ Treatment + (1|colony) + (1|inoc_date), family = binomial(link = "logit"), data = gris_alive)
gris_presence5 <- glmer(Presence ~ Treatment + (1|colony) , family = binomial(link = "logit"), data = gris_alive)

AICtab(gris_presence, gris_presence2, gris_presence3, gris_presence4, gris_presence5) # models 5 and 2 do best 
Anova(gris_presence5)
Anova(gris_presence2)
summary(gris_presence)
Anova(gris_presence)
anova(gris_presence, test = "Chisq")
```

Use emmeans since predict function does not work with mixed effect models: 

```{r}
emmeans(gris_presence5,~Treatment, type = "response") 
```


### Mortality 
#### With just SS  
```{r}
# make a survival object
event <- gris2$dead
time <- gris2$days_til_dead
survival <- Surv(time, event)
```

```{r}
# Models with just SS 

mortg1<- coxph(survival ~ Treatment + MCL, data=gris2) 
mortg2<- coxme(survival ~ Treatment + MCL + (1|inoc_date), data=gris2)
mortg3 <- coxme(survival ~ Treatment + (1|inoc_date), data = gris2)

AICtab(mortg1, mortg2, mortg3)
```

```{r}
Anova(mortg2)
summary(glht(mortg2,linfct=mcp(Treatment="Tukey")))
```

#### With all strains 

```{r}
# make a survival object
event <- gris$dead
time <- gris$days_til_dead
survival <- Surv(time, event)
```

```{r}
# model survival as the response with the cox proportional hazards model 
# be careful because crithidia strain is confounded by inoculation date 

mortg1<- coxph(survival ~ Treatment * crithidia_strain + MCL, data=gris) 
mortg2<- coxph(survival ~ Treatment + crithidia_strain + MCL, data=gris) 

mortg3<- coxme(survival ~ Treatment * crithidia_strain + MCL + (1|inoc_date), data=gris)
mortg4<- coxme(survival ~ Treatment + crithidia_strain + MCL + (1|inoc_date), data=gris)

mortg5<- coxph(survival ~ Treatment + MCL, data=gris)
mortg6<- coxph(survival ~ crithidia_strain + MCL, data=gris)

mortg7<- coxph(survival ~ Treatment, data=gris)
mortg8<- coxph(survival ~ crithidia_strain, data=gris)

mortg9<- coxme(survival ~ Treatment + MCL + (1|inoc_date), data=gris)
mortg10<- coxme(survival ~ crithidia_strain + MCL + (1|inoc_date), data=gris)

mortg11<-coxph(survival ~ Treatment + crithidia_strain + MCL + inoc_date, data = gris)


AICtab(mortg1, mortg2, mortg3, mortg4, mortg5, mortg6, mortg7, mortg8, mortg9, mortg10, mortg11)
```

best models exclude treatment. but we'll use model #4 which includes treatment because that's our term of interest. 

```{r}
ggcoxdiagnostics(mortg11, 
                 type = "deviance",
                 linear.predictions = FALSE, 
                 ggtheme = theme_bw())  
# just to see - not the model we're using! 
```

```{r}
Anova(mortg4)
```

```{r}
summary(glht(mortg4,linfct=mcp(crithidia_strain="Tukey")))
```
