---
title: "2018 Count Data Analyses"
output: 
  html_document:
    toc: true
    toc_depth: 3
    highlight: pygments
---

This R code is for analyzing the effect of pollen diet on Crithidia infection in wild Bombus griseocollis and Bombus bimaculatus workers for the manuscript titled: "Sunflower pollen reduces a gut pathogen in the model bee species, Bombus impatiens, but has weaker effects in three wild congeners"

Authors: Alison Fowler*, Jonathan Giacomini, Sara June Connon, Becky Irwin, and Lynn Adler 
Corresponding author*: aefowler@umass.edu 

## Libraries 
```{r, results = 'hide', message = FALSE, warning = FALSE}
library(lattice)
library(psych)
library(car)
library(AER)
library(MASS)
library(corrplot)
library(AICcmodavg)
library(dplyr)
library(tidyr)
library(lme4)
library(mgcv)
library(scatterplot3d)
library(bbmle)
library(emmeans)
library(ggplot2)
library(tidyverse)
library(ggalt)
library(multcomp)
library(DHARMa)
```

## *B. bimaculatus* 

### Get set up 

```{r, message = FALSE}
bimacs18<- read_csv("CSV/2018_bimacs.csv")

bimacs18 <- bimacs18 %>% 
  dplyr::rename(count = "Crithidia Count") %>% 
  dplyr::rename(ID = "Bee ID") %>% 
  dplyr::rename(inoc_date = "Inoc. Date") %>% 
  dplyr::rename(crithidia_strain = "Crithidia Strain") %>% 
  dplyr::rename(colony = "Colony ID")

# remove the bees that didn't finish inoculum (48, 95, 50) and only  bee from colony 112 (49)
bimacs18 <- bimacs18 %>% 
  filter(ID != "48") %>% 
  filter(ID != "95") %>% 
  filter(ID != "50") %>% 
  filter(ID != "49")

# make colony ID a factor 
bimacs18$colony<-as.factor(bimacs18$colony)

# make treat a factor 
bimacs18$Treatment<-as.factor(bimacs18$Treatment)

# make Crithidia strain a factor
bimacs18$crithidia_strain <- as.factor(bimacs18$crithidia_strain )

# make inoc date a factor
bimacs18$inoc_date <- as.factor(bimacs18$inoc_date)

# add column for presence/absence 
bimacs18$Presence <- NA

for(i in 1:length(bimacs18$count)){
  if(bimacs18$count[i] >=1) {
    bimacs18$Presence[i] <- 1
  } else {
    bimacs18$Presence[i] <- 0
  }
} 


class(bimacs18$crithidia_strain)
# change order of the strains to make the plot easier 
levels(bimacs18$crithidia_strain)
bimacs18$crithidia_strain <- relevel(bimacs18$crithidia_strain, "170-B (Colony 170)")
bimacs18$crithidia_strain <- relevel(bimacs18$crithidia_strain, "129-B, W3-B (Colony 1)")
levels(bimacs18$crithidia_strain)
```

### Data exploration 

```{r}
summary(bimacs18$inoc_date)
summary(bimacs18$colony)
summary(bimacs18$Treatment)
summary(bimacs18$crithidia_strain)

boxplot(count ~ inoc_date, data=bimacs18)
boxplot(count ~ Treatment, data=bimacs18)
boxplot(count ~ colony, data=bimacs18)
boxplot(count ~ crithidia_strain, data=bimacs18)

# look at the raw means that correspond with those boxplots 

tapply(bimacs18$count, list(bimacs18$Treatment), mean)
tapply(bimacs18$count, list(bimacs18$colony), mean) 
tapply(bimacs18$count, list(bimacs18$inoc_date), mean)
tapply(bimacs18$count, list(bimacs18$crithidia_strain), mean)

# plot coloreed by colony (ggalt package)
(ggplot(bimacs18, aes(x=Treatment, y=count, shape=colony))
  +geom_point(aes(color=colony),alpha=0.5)
  +geom_encircle(aes(color=colony, group=colony))
  +theme(legend.position="left"))

# plot colored by Crithidia source
(ggplot(bimacs18, aes(x=Treatment, y=count, shape=crithidia_strain))
  +geom_point(aes(color=crithidia_strain),alpha=0.5)
  +geom_encircle(aes(color=crithidia_strain, group=crithidia_strain))
  +theme(legend.position="left"))

hist(bimacs18$count, breaks = 20)

look <- c("count", "MCL", "colony", "inoc_date", "Treatment")
pairs.panels(bimacs18[,look], scale= TRUE, method="spearman")

plot(count ~ MCL, data=bimacs18)
```

### Model Infection Intensity 

```{r}
# strain as interaction
bmod1<-glmer.nb(count ~ Treatment*crithidia_strain + MCL + (1|colony) + (1|inoc_date), data = bimacs18)
isSingular(bmod1)

# one RE
bmod2<-glmer.nb(count ~ Treatment*crithidia_strain + MCL + (1|colony) , data = bimacs18)
bmod3<-glmer.nb(count ~ Treatment*crithidia_strain + MCL + (1|inoc_date), data = bimacs18)
isSingular(bmod3)

# no RE but include wing
bmod4<-glm.nb(count ~ Treatment*crithidia_strain + MCL , data = bimacs18)

# REs but no wing 
bmod5<-glmer.nb(count ~ Treatment*crithidia_strain + (1|colony) + (1|inoc_date), data = bimacs18)
isSingular(bmod5)
bmod6<-glmer.nb(count ~ Treatment*crithidia_strain + (1|colony), data = bimacs18)
bmod7<-glmer.nb(count ~ Treatment*crithidia_strain + (1|inoc_date), data = bimacs18)
isSingular(bmod7)

# neither wing nor REs
bmod8<-glm.nb(count ~ Treatment*crithidia_strain, data = bimacs18)

# strain as additive 
# with wing
bmod9<-glmer.nb(count ~ Treatment + crithidia_strain + MCL + (1|colony) + (1|inoc_date), data = bimacs18)
isSingular(bmod9)
bmod10<-glmer.nb(count ~ Treatment + crithidia_strain + MCL + (1|colony) , data = bimacs18)
bmod11<-glmer.nb(count ~ Treatment + crithidia_strain + MCL + (1|inoc_date), data = bimacs18)
isSingular(bmod11)
bmod12<-glm.nb(count ~ Treatment + crithidia_strain + MCL , data = bimacs18)

#without wing
bmod13<-glmer.nb(count ~ Treatment + crithidia_strain + (1|colony) + (1|inoc_date), data = bimacs18)
isSingular(bmod13)
bmod14<-glmer.nb(count ~ Treatment + crithidia_strain + (1|colony), data = bimacs18)
bmod15<-glmer.nb(count ~ Treatment + crithidia_strain + (1|inoc_date), data = bimacs18)
isSingular(bmod15)
bmod16<- glm.nb(count ~ Treatment + crithidia_strain, data = bimacs18)

AICtab(bmod1,bmod2,bmod3,bmod4,bmod5,bmod6,bmod7,bmod8,bmod9,bmod10,bmod11,bmod12,bmod13,bmod14,bmod15,bmod16)
```

Model for 2018 bimaculatus: 
Model 12: count ~ Treatment + crithidia_strain + MCL 

```{r}
# look at model fit 
par(mfrow=c(2,2))
plot(bmod12)
simresid <- simulateResiduals(bmod12, plot = T)

# look at output 
Anova(bmod12)
summary(bmod12)
```

### Plot Intensity

```{r}
bimac_18_means<-emmeans(bmod12, ~Treatment + crithidia_strain) 
b.18.means.to.plot<-as.data.frame(summary(bimac_18_means))
b.18.means.to.plot

# exponentiate to get back on response scale
b.18.means.to.plot$expmean<- exp(b.18.means.to.plot$emmean)
b.18.means.to.plot$tfupper<-exp(b.18.means.to.plot$emmean + b.18.means.to.plot$SE)
b.18.means.to.plot$tlower<-exp(b.18.means.to.plot$emmean - b.18.means.to.plot$SE)
exp.mean<-b.18.means.to.plot$expmean
b.18.means.to.plot
```

```{r}
# create letters from pair-wise tests 
bimac.18.lets<-cld(bimac_18_means, Letters = c('a','b','c', 'd'), sort = FALSE)
bimac.18.letters<-dplyr::select(bimac.18.lets, Treatment, crithidia_strain, .group)
bimac.18.letters
```

```{r}
bim_2018_plot<-
  ggplot(b.18.means.to.plot, aes(x=Treatment, group = crithidia_strain,
                          y=b.18.means.to.plot$expmean, 
                          ymin=b.18.means.to.plot$tlower,
                          ymax=b.18.means.to.plot$tfupper)) + 
  theme_classic()+
  geom_point(stat="identity",
             aes(shape = crithidia_strain,
                 color = Treatment),
             position = position_dodge(width = 0.2),
             size = 5.5) +
  geom_errorbar(width=0.05, 
                aes(color = Treatment),
                position = position_dodge(width = 0.2))+
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"))+
  theme(text = element_text(size=18))+
  scale_shape_manual(expression(paste(italic("Crithidia"))),
                     labels = c("BIM-2",
                                "BIM-1",
                                "BIM-3"),
                     values = c(15, 17, 19),
                     guide = "legend") + 
  scale_color_manual(values = c("chocolate2", "royalblue3")) + 
  guides(color = FALSE) +
  theme(legend.position = c(.5,.7)) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.title.y = element_blank()) + 
  ggtitle(expression(paste(italic("B. bimaculatus ")))) + 
  geom_text(x = 2.35, y = 230, label = "2018", size = 4)
#  theme(plot.title = element_text(hjust = 0.5))
```

### Presence/Absence 
```{r}
# no REs 
bimacs_18_presence_add <- glm(Presence ~ Treatment + crithidia_strain + MCL, family = binomial(link = "logit"), data = bimacs18)

bimacs_18_presence_int <- glm(Presence ~ Treatment * crithidia_strain + MCL, family = binomial(link = "logit"), data = bimacs18)

bimacs_18_presence_wing <- glm(Presence ~ Treatment + MCL, family = binomial(link = "logit"), data = bimacs18)

bimacs_18_presence_no_wing <- glm(Presence ~ Treatment, family = binomial(link = "logit"), data = bimacs18)

# int models with REs 
bimacs_18_presence_int_both <- glmer(Presence ~ Treatment * crithidia_strain + MCL + (1|colony) + (1|inoc_date), family = binomial(link = "logit"), data = bimacs18) # failed to converge 

bimacs_18_presence_int_inoc <- glmer(Presence ~ Treatment * crithidia_strain + MCL + (1|inoc_date), family = binomial(link = "logit"), data = bimacs18)

bimacs_18_presence_int_col <- glmer(Presence ~ Treatment * crithidia_strain + MCL + (1|colony), family = binomial(link = "logit"), data = bimacs18)

# models without source but with REs
bimacs_18_presence_wing_both <- glmer(Presence ~ Treatment + MCL + (1|colony) + (1|inoc_date), family = binomial(link = "logit"), data = bimacs18)

bimacs_18_presence_wing_inoc <- glmer(Presence ~ Treatment + MCL + (1|inoc_date), family = binomial(link = "logit"), data = bimacs18)

bimacs_18_presence_wing_col <- glmer(Presence ~ Treatment + MCL + (1|colony), family = binomial(link = "logit"), data = bimacs18)

bimacs_18_presence_nowing_col <- glmer(Presence ~ Treatment + (1|colony), family = binomial(link = "logit"), data = bimacs18)

AICtab(bimacs_18_presence_int_both, 
       bimacs_18_presence_int_col, 
       bimacs_18_presence_int_inoc, 
       bimacs_18_presence_int, 
       bimacs_18_presence_wing, 
       bimacs_18_presence_wing_both, 
       bimacs_18_presence_wing_inoc, 
       bimacs_18_presence_wing_col, 
       bimacs_18_presence_add, 
       bimacs_18_presence_no_wing,
       bimacs_18_presence_nowing_col)

# wing col does best
```

```{r}
# look at significance of colony: 

anova(bimacs_18_presence_int_col, bimacs_18_presence_int) 
# colony is highly significant with interaction term p = 0.004 

anova(bimacs_18_presence_wing_col, bimacs_18_presence_wing) 
# still significant but not as much without the interaction term p = 0.011
```

```{r}
# look at outputs from model with colony as random effect (and no source term)
Anova(bimacs_18_presence_wing_col)
```

```{r}
# look at outputs from model with colony and source interaction 
Anova(bimacs_18_presence_int_col)

summary(bimacs_18_presence_int_col)

emmeans(bimacs_18_presence_int_col, specs = pairwise ~ Treatment|crithidia_strain, type = "response")
```

```{r}
# look at means of interaction model
bimac_18_p_int_means<-emmeans(bimacs_18_presence_int, ~Treatment * crithidia_strain, type = "response")

# look at tukey's adjusted pairwise comparisons
bimac.18.p.lets<-cld(bimac_18_p_int_means, Letters = c('a','b','c', 'd'), sort = FALSE)
```

Model we're going to use: 
bimacs_18_presence_wing_col

```{r}
summary(bimacs_18_presence_wing_col)
Anova(bimacs_18_presence_wing_col)

emmeans(bimacs_18_presence_wing_col, ~Treatment, type = "response")
```

### Mortality 

```{r}
bimacsfull<-read_csv("CSV/bimaculatus_full_data.csv")

bimacsfull <- bimacsfull %>% 
  rename(ID = "Bee ID") %>% 
  rename(inoc_date = "Inoc. Date") %>% 
  rename(crithidia_strain = "Crithidia Strain") %>% 
  rename(colony = "Colony ID") %>% 
  rename(MCL = "MCL-MM")

# remove the bees that didn't finish inoculum (48, 95, 50, 28) and only  bee from colony 112 (49)
bimacsfull <- bimacsfull %>% 
  filter(ID != "48") %>% 
  filter(ID != "95") %>% 
  filter(ID != "50") %>% 
  filter(ID != "49") %>% 
  filter(ID != "28")

# make colony ID a factor 
bimacsfull$colony<-as.factor(bimacsfull$colony)

# make treat a factor 
bimacsfull$Treatment<-as.factor(bimacsfull$Treatment)

# make Crithidia strain a factor
bimacsfull$crithidia_strain <- as.factor(bimacsfull$crithidia_strain )

# make inoc date a factor
bimacsfull$inoc_date <- as.factor(bimacsfull$inoc_date)

summary(bimacsfull$dead)
```

```{r}
# make a survival object
event <- bimacsfull$dead
time <- bimacsfull$days_til_dead
survival <- Surv(time, event)
```

```{r}
# model survival as the response with the cox proportional hazards model 

# don't have wing measurements for dead bees so exclude wing

b1<- coxph(survival ~ Treatment * crithidia_strain, data=bimacsfull) 
b2<- coxph(survival ~ Treatment + crithidia_strain, data=bimacsfull) 
b3<- coxph(survival ~ Treatment, data=bimacsfull)
b4<- coxph(survival ~ crithidia_strain, data=bimacsfull)

b5<- coxme(survival ~ Treatment * crithidia_strain + (1|inoc_date), data=bimacsfull)
b6<- coxme(survival ~ Treatment + crithidia_strain + (1|inoc_date), data=bimacsfull)
b7<- coxme(survival ~ crithidia_strain + (1|inoc_date), data=bimacsfull)
b8<- coxme(survival ~ Treatment + (1|inoc_date), data=bimacsfull)

b9<-coxph(survival ~ Treatment + crithidia_strain + inoc_date, data = bimacsfull)


AICtab(b1, b2, b3, b4, b5, b6, b7, b8, b9)
```

best models didn't include treatment. but we'll use b8. 

```{r}
ggcoxdiagnostics(b2, 
                 type = "deviance",
                 linear.predictions = FALSE, 
                 ggtheme = theme_bw())  
# just to see - not the model we're using! 
```

```{r}
Anova(b8)
```

Plots

```{r}
mortb_plot <- survfit(survival ~ Treatment, data=bimacsfull)
ggsurvplot(mortb_plot, legend = "left",palette = c("orange","red","forestgreen"))
```


## *B. vagans* 

### Get set up 

```{r}
vagans<- read_csv("CSV/2018_vagans.csv")

vagans <- vagans %>% 
  dplyr::rename(ID = "Bee ID") %>% 
  dplyr::rename(inoc_date = "Inoc. Date") %>% 
  dplyr::rename(crithidia_strain = "Crithidia Strain") %>% 
  dplyr::rename(colony = "Colony ID") %>% 
  dplyr::rename(wing = "MCL-MM")

# remove the bees that didn't finish inoculum (108, 112) and escapes (104, 162) and dead (98)
vagans <- vagans %>% 
  filter(ID != "108") %>% 
  filter(ID != "112") %>% 
  filter(ID != "104") %>% 
  filter(ID != "98") %>% 
  filter(ID != "162")

# make colony ID a factor 
vagans$colony<-as.factor(vagans$colony)

# make treat a factor 
vagans$Diet<-as.factor(vagans$Diet)

# make Crithidia strain a factor
vagans$crithidia_strain <- as.factor(vagans$crithidia_strain )

# make inoc date a factor
vagans$inoc_date <- as.factor(vagans$inoc_date)

# replace lost wing values (NAs) with mean wing 
vagans<- vagans %>% 
  mutate(wing2 = case_when(is.na(wing) ~ mean(wing, na.rm = TRUE), 
                          TRUE ~ as.numeric(wing)
                          )
         )

# add column for presence/absence 
vagans$Presence <- NA

for(i in 1:length(vagans$Count)){
  if(vagans$Count[i] >=1) {
    vagans$Presence[i] <- 1
  } else {
    vagans$Presence[i] <- 0
  }
} 

```

### Data exploration 

```{r}
summary(vagans$inoc_date)
summary(vagans$Diet)

boxplot(Count ~ inoc_date, data=vagans)
boxplot(Count ~ Diet, data=vagans)

# look at the raw means that correspond with those boxplots 

tapply(vagans$Count, list(vagans$Diet), mean)
tapply(vagans$Count, list(vagans$inoc_date), mean)

hist(vagans$Count, breaks = 20)

look <- c("Count", "wing2", "inoc_date", "Diet")
pairs.panels(vagans[,look], scale= TRUE, method="spearman")

plot(Count ~ wing2, data=vagans)
```

### Model Intensity

```{r}
vmod1<-glmer.nb(Count ~ Diet + wing2 + (1|inoc_date), data = vagans)
vmod2<-glmer.nb(Count ~ Diet + (1|inoc_date), data = vagans)
vmod3<-glm.nb(Count ~ Diet + wing2, data = vagans)
vmod4<-glm.nb(Count ~ Diet, data = vagans)

AICtab(vmod1, vmod2, vmod3, vmod4)
```

```{r}
plot(vmod2)
simresid <- simulateResiduals(vmod2, plot = T)
Anova(vmod2)
```

### Plot Intensity

```{r}
vmeans<-emmeans(vmod2, ~Diet) 
v.means.to.plot<-as.data.frame(summary(vmeans))
v.means.to.plot$expmean<- exp(v.means.to.plot$emmean)
v.means.to.plot$tfupper<-exp(v.means.to.plot$emmean + v.means.to.plot$SE)
v.means.to.plot$tlower<-exp(v.means.to.plot$emmean - v.means.to.plot$SE)
exp.mean<-v.means.to.plot$expmean
v.means.to.plot

v.lets<-cld(vmeans, Letters = c('a','b','c', 'd'), sort = FALSE)
vletters<-dplyr::select(v.lets, Diet, .group)
```

```{r}
vagans_plot<-
  ggplot(v.means.to.plot, aes(x=Diet, group = Diet,
                          y=v.means.to.plot$expmean, 
                          ymin=v.means.to.plot$tlower,
                          ymax=v.means.to.plot$tfupper)) + 
  theme_classic()+
  geom_point(stat="identity",
             aes(color = Diet),
             size = 5.5) +
  geom_errorbar(width=0.05, aes(color = Diet)) +
  theme(axis.text.x = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"), 
        axis.text.y = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"))+
  theme(text = element_text(size=18))+
  scale_color_manual(values = c("chocolate2", "royalblue3")) + 
  scale_x_discrete(labels = c("100% SF", "Control")) +
  guides(color = FALSE) +
  theme(axis.title.x = element_blank()) + 
  theme(axis.title.y = element_blank()) + 
  ggtitle(expression(paste(italic("B. vagans*")))) + 
  geom_text(x=2.35, y=57.5, label="2018", size = 4) 

vagans_plot
```

### Presence/Absence for vagans

```{r}
vagans_presence <- glm(Presence ~ Diet, family = binomial(link = "logit"), data = vagans)
summary(vagans_presence)
Anova(vagans_presence)
anova(vagans_presence, test = "Chisq")
```

```{r}
Diet <- c("SF", "WF")
newdata <- data.frame(Diet)
newdata$dietP <- predict(vagans_presence, newdata = newdata, type = "response")
newdata$inverse <- 1-newdata$dietP
newdata
```

```{r}
newdata2 <- cbind(newdata, predict(vagans_presence, newdata = newdata, type = "link", se = T))
newdata2
```

```{r}
newdata3 <- within(newdata2, {
    PredictedProb <- plogis(fit)
    LL <- plogis(fit - (1.96 * se.fit)) # this translates to a 95% CI  
    UL <- plogis(fit + (1.96 * se.fit))
})

newdata3
```


### No mortality analysis ; only one bee died (SF). 