---
title: "2020_imps"
output: 
  html_document:
    toc: true
    toc_depth: 3
    highlight: pygments
---

This R code is for analyzing the effect of pollen diet on Crithidia infection in wild Bombus impatiens workers for the manuscript titled: "Sunflower pollen reduces a gut pathogen in the model bee species, Bombus impatiens, but has weaker effects in three wild congeners"

Authors: Alison Fowler*, Jonathan Giacomini, Sara June Connon, Becky Irwin, and Lynn Adler 
Corresponding author*: aefowler@umass.edu 

## Libraries 

```{r, results = 'hide', message = FALSE, warning = FALSE}
library(lattice)
library(psych)
library(car)
library(AER)
library(MASS)
library(corrplot)
library(AICcmodavg)
library(dplyr)
library(tidyr)
library(lme4)
library(mgcv)
library(scatterplot3d)
library(bbmle)
library(emmeans)
library(ggplot2)
library(tidyverse)
library(ggalt)
library(multcomp)
library(DHARMa)
```

## Get set up 

```{r, message = F}
imps20<- read_csv("CSV/2020_imp_counts.csv")

imps20<- imps20 %>% 
  dplyr::rename(Bee_ID = 'Bee ID') %>% 
  dplyr::rename(Colony = 'Colony ID') %>% 
  dplyr::rename(Date = "Date Inoc")

# make treatment a factor
imps20$Treatment <- as.factor(imps20$Treatment)

# make colony ID a factor 
imps20$Colony<-as.factor(imps20$Colony)

# make date a factor 
imps20$Date<-as.factor(imps20$Date)

# reorder treatments 
imps20$Treatment <- factor(imps20$Treatment, levels=c("SUN", "MIX", "WF"))

#levels(imps20$Treatment)

# rename treatments 
levels(imps20$Treatment)[levels(imps20$Treatment) =="SUN"] <- "100% Sun"
levels(imps20$Treatment)[levels(imps20$Treatment) =="MIX"] <- "50% Sun"
levels(imps20$Treatment)[levels(imps20$Treatment) =="WF"] <- "Wildflower mix"

# remove dead bees for count analysis   
imps20<-imps20 %>% 
  filter(Dead != '1')

# replace lost wing values (NAs) with mean wing 
imps20<- imps20 %>% 
  mutate(wing = case_when(is.na(Wing_mm) ~ mean(Wing_mm, na.rm = TRUE), 
                          TRUE ~ as.numeric(Wing_mm)
                          )
         )

# add column for presence/absence 
imps20$Presence <- NA

for(i in 1:length(imps20$Count)){
  if(imps20$Count[i] >=1) {
    imps20$Presence[i] <- 1
  } else {
    imps20$Presence[i] <- 0
  }
}
```

## Counts 
### Data exploration 

```{r}
summary(imps20$Date)
summary(imps20$Colony)
summary(imps20$Treatment)

# counts by treatments 
imps20 %>% 
  ggplot(aes(x = Treatment, y = Count)) + 
  geom_boxplot() + 
  geom_point(alpha = 0.5)

# counts by colonies 
imps20 %>% 
  ggplot(aes(x = Treatment, y = Count)) + 
  geom_point(aes(color = Colony))

imps20 %>% 
  ggplot(aes(x = Colony, y = Count)) + 
  geom_point(aes(color = Treatment))

imps20 %>% 
  ggplot(aes(x = Colony, y = Count)) + 
  geom_boxplot() +
  geom_point(aes(color = Treatment))

imps20 %>% 
  ggplot(aes(x = Date, y = Count)) + 
  geom_boxplot()+
  geom_point(aes(color = Treatment))

# distribution of counts
hist(imps20$Count)
```

### Intensity models  

```{r}
imp1<- glmer.nb(Count ~ Treatment + wing + (1|Colony) + (1|Date), data = imps20)
imp2<- glmer.nb(Count ~ Treatment + wing + (1|Date), data = imps20)
imp3<- glmer.nb(Count ~ Treatment + wing + (1|Colony), data = imps20)
imp4<- glm.nb(Count ~ Treatment + wing, data = imps20)

imp5<- glmer.nb(Count ~ Treatment + (1|Colony) + (1|Date), data = imps20)
imp6<- glmer.nb(Count ~ Treatment + (1|Date), data = imps20)
imp7<- glmer.nb(Count ~ Treatment + (1|Colony) , data = imps20)
imp8<- glm.nb(Count ~ Treatment, data = imps20)

AICtab(imp1,imp2,imp3,imp4,imp5,imp6,imp7,imp8)
```

```{r}
plot(imp7)
plot(imp3)
plot(imp5)

simresid <- simulateResiduals(imp7, plot = T)

Anova(imp7)
```

### Intensity Plot

```{r}
imp_means<-emmeans(imp7, ~Treatment) 
imp.20.means.to.plot<-as.data.frame(summary(imp_means))
imp.20.means.to.plot$expmean<- exp(imp.20.means.to.plot$emmean)
imp.20.means.to.plot$tfupper<-exp(imp.20.means.to.plot$emmean + imp.20.means.to.plot$SE)
imp.20.means.to.plot$tlower<-exp(imp.20.means.to.plot$emmean - imp.20.means.to.plot$SE)
exp.mean<-imp.20.means.to.plot$expmean
imp.20.means.to.plot

imp.lets<-cld(imp_means, Letters = c('a','b','c', 'd'), sort = FALSE)
imp.letters<-dplyr::select(imp.lets, Treatment, .group)
```

```{r}
imp_2020_plot<-
  ggplot(imp.20.means.to.plot, aes(x=Treatment, group = Treatment,
                          y=imp.20.means.to.plot$expmean, 
                          ymin=imp.20.means.to.plot$tlower,
                          ymax=imp.20.means.to.plot$tfupper)) + 
  theme_classic()+
  geom_point(stat="identity",
             aes(color = Treatment),
             size = 5) +
  geom_errorbar(width=0.05, aes(color = Treatment))+
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_text(face="plain",
                                   size = 15,
                                   angle = 0,
                                   vjust = 0.5, 
                                   hjust = 0.5,
                                   colour = "black"))+
  theme(text = element_text(size=15))+
  geom_text(aes (y = imp.20.means.to.plot$expmean, 
                 label = imp.letters$.group), 
            size = 5, hjust = 1.5) +
  scale_color_manual(values = c("chocolate2", "goldenrod1", "royalblue3")) + 
  guides(color = FALSE) +
  theme(axis.title.x = element_blank()) + 
  theme(axis.title.y = element_blank()) + 
  ggtitle("")+ 
  geom_text(x = 3.35, y = 77.5, label = "2020", size = 4) 
```

### Presence/Absence models 

```{r}
imps_presence <- glmer(Presence ~ Treatment + (1|Colony), family = binomial(link = "logit"), data = imps20)
imps_presence2 <- glm(Presence ~ Treatment, family = binomial(link = "logit"), data = imps20)

AICtab(imps_presence, imps_presence2)

summary(imps_presence2)
Anova(imps_presence2)
```

```{r}
Treatment <- c("100% Sun", "50% Sun", "Wildflower mix")
newdata <- data.frame(Treatment)
newdata$dietP <- predict(imps_presence2, newdata = newdata, type = "response")
newdata$inverse <- 1-newdata$dietP
newdata
```

```{r}
newdata2 <- cbind(newdata, predict(imps_presence2, newdata = newdata, type = "link", se = T))
newdata2
```

```{r}
newdata3 <- within(newdata2, {
    PredictedProb <- plogis(fit)
    LL <- plogis(fit - (1.96 * se.fit)) # this translates to a 95% CI  
    UL <- plogis(fit + (1.96 * se.fit))
})

newdata3
```

## Mortality 

```{r}
# make a survival object
event <- imps20$Dead
time <- imps20$days_til_dead
survival <- Surv(time, event)
```

```{r}
# model survival as the response with the cox proportional hazards model 

mimp1<- coxme(survival ~ Treatment + wing + (1|Colony) + (1|Date), data = imps20)
mimp2<- coxme(survival ~ Treatment + wing + (1|Date), data = imps20)
mimp3<- coxme(survival ~ Treatment + wing + (1|Colony), data = imps20)
mimp4<- coxph(survival ~ Treatment + wing, data = imps20)

mimp5<- coxme(survival ~ Treatment + (1|Colony) + (1|Date), data = imps20)
mimp6<- coxme(survival ~ Treatment + (1|Date), data = imps20)
mimp7<- coxme(survival ~ Treatment + (1|Colony) , data = imps20)
mimp8<- coxph(survival ~ Treatment, data = imps20)

mimp9<- coxph(survival ~ Treatment + wing + Colony, data=imps20)
mimp10<- coxph(survival ~ Treatment + Colony, data=imps20)

AICtab(mimp1,mimp2,mimp3,mimp4,mimp5,mimp6,mimp7,mimp8, mimp9, mimp10)
```

```{r}
Anova(mimp7)
```

## Consumption

did not measure evaporation

```{r, message = F}
imp_cons <- read_csv("CSV/wild_imps_2020_consumption.csv")

imp_cons <- imp_cons %>% 
  rename(New_pollen = `New pollen weight`) %>% 
  rename(Old_pollen = `Old pollen weight`) %>% 
  rename(New_sucrose = `New sucrose weight`) %>% 
  rename(Old_sucrose = `Old sucrose weight`) %>% 
  rename(Net_pollen = `Net pollen`) %>% 
  rename(Net_sucrose = `Net sucrose`) %>% 
  rename(Bee_ID = `Bee ID`)

# remove bee # 56; it died 
imp_cons <- imp_cons %>% 
  filter(Bee_ID != "56")
```

```{r}
imps_20_wings <- data.frame(imps20$Bee_ID, imps20$wing, imps20$Count)
imps_20_wings <- imps_20_wings %>% 
  rename(Bee_ID = imps20.Bee_ID) %>% 
  rename(Wing = imps20.wing) %>% 
  rename(Count = imps20.Count)
```

```{r}
full_imp_cons <- merge(imp_cons, imps_20_wings, by = "Bee_ID")

full_imp_cons <- full_imp_cons %>% 
  rename(Colony_ID = `Colony ID`) 

full_imp_cons$Treatment <- as.factor(full_imp_cons$Treatment)
full_imp_cons$Colony_ID <- as.factor(full_imp_cons$Colony_ID)

# rename diets 
levels(full_imp_cons$Treatment)[levels(full_imp_cons$Treatment) =="MIX"] <- "50% Sunflower"
levels(full_imp_cons$Treatment)[levels(full_imp_cons$Treatment) =="WF"] <- "Wildflower"
levels(full_imp_cons$Treatment)[levels(full_imp_cons$Treatment) =="SUN"] <- "100% Sunflower"
```

```{r}
boxplot(Net_pollen ~ Treatment, data = full_imp_cons)
boxplot(Net_pollen ~ Colony_ID, data = full_imp_cons)
boxplot(Net_sucrose ~ Treatment, data = full_imp_cons)
boxplot(Net_pollen ~ `New date`, data = full_imp_cons)
boxplot(Net_sucrose ~ `New date`, data = full_imp_cons)
plot(Net_pollen ~ Wing, data = full_imp_cons)
plot(Net_sucrose ~ Wing, data = full_imp_cons)
hist(full_imp_cons$Net_pollen)
hist(full_imp_cons$Net_sucrose)

plot(Net_pollen~ Count, data = full_imp_cons)
plot(Count~ Net_pollen, data = full_imp_cons, col = Treatment)

figS3<- 
  ggplot(full_imp_cons, aes(x = Net_pollen, y = Count,color = Treatment)) + 
  geom_point(aes(color = Treatment), size = 4) + 
  geom_smooth(method='lm', se = F, size = 1) +
  xlab("Pollen consumed (g)") + 
  ylab("Count") + 
  theme_bw() + 
  theme(text = element_text(size=18)) + 
  scale_color_manual(values = c("goldenrod1", "chocolate2", "royalblue3"))

ggplot(full_imp_cons, aes(x = Net_pollen, y = Count)) + 
  geom_point(aes(color = Treatment)) + 
  geom_smooth(method='lm', se = F, color = "black") +
  xlab("Pollen consumed (g)") + 
  ylab("Count") 
```

### Models 

#### Pollen 
```{r}
imps_p_mod <- lmer(Net_pollen ~ Treatment + Wing + (1|Colony_ID) + (1|`New date`), data = full_imp_cons)
imps_p_mod2 <- lm(Net_pollen ~ Treatment + Wing, data = full_imp_cons)
imps_p_mod3 <- lmer(Net_pollen ~ Treatment + (1|Colony_ID), data = full_imp_cons)
imps_p_mod4 <- lmer(Net_pollen ~ Treatment + (1|`New date`), data = full_imp_cons)
imps_p_mod5 <- lm(Net_pollen ~ Treatment, data = full_imp_cons)

AICtab(imps_p_mod, imps_p_mod2, imps_p_mod3, imps_p_mod4, imps_p_mod5)
```

```{r}
# diagnostics 
plot(imps_p_mod5)
simresid <- simulateResiduals(imps_p_mod5, plot = T)

# outputs 
summary(imps_p_mod5)
Anova(imps_p_mod5)
summary(glht(imps_p_mod5,linfct=mcp(Treatment="Tukey")))
```

#### Sucrose 
```{r}
imps_s_mod <- lmer(Net_sucrose ~ Treatment + Wing_mm + (1|Colony_ID), data = full_imp_cons)
imps_s_mod2 <- lm(Net_sucrose ~ Treatment + Wing_mm, data = full_imp_cons)
imps_s_mod3 <- lmer(Net_sucrose ~ Treatment + (1|Colony_ID), data = full_imp_cons)
imps_s_mod4 <- lm(Net_sucrose ~ Treatment, data = full_imp_cons)

AICtab(imps_s_mod, imps_s_mod2, imps_s_mod3, imps_s_mod4)
```

```{r}
plot(imps_s_mod2)
simresid <- simulateResiduals(imps_s_mod2, plot = T)

summary(imps_s_mod2)
Anova(imps_s_mod2)
```

```{r}
pollen_count_mod <- glm.nb(Count ~ Net_pollen, data = full_imp_cons)

plot(pollen_count_mod)
simresid <- simulateResiduals(pollen_count_mod, plot = T)
summary(pollen_count_mod)
```

Pollen treatment is significant. Sucrose non-significant. 
